Kindly Bent to Free Us

### 2019-11-11 kindly/infer/infer.tex
Fixed uses of 'note' instead of 'denote' and various typos.

Unlike the syntax-directed version of our type system, knowing which elements
are input and output or our inference judgement is critical. In the rest
of this section, when presenting a new judgement
we will note in \textbf{bold} the input parameters. The rest will then be
output parameters.

@@

% One novelty of our inference  judgement now
% returns a type environment, which we call ``usage environment'' and commonly
% note $\Sv$, to summarize how variables and borrows are used inside
% the expression.

In order to determine if a variable is used in an affine manner, we must track
its uses and the associated kinds. For instance, in the expression
$(x,x)$, $x$ is used twice. If $x$ is of type $\tau$, which is of kind $k$,
we must add the constraint $\Cleq{k}{\kun}$.
%
In order to infer such constraints, our inference judgement will not only
take an environment as parameter but also return an environment which
we call ``usage environment'' and which summarizes how variables and borrows
are used. Usage environments follow the exact same grammar
as normal environment. In order to distinguish them more easily,
we note them $\Sv$.

@@

We note $\inferW{\Sv}{(C,\unif)}{\bm{\E}}{\bm{e}}{\tau}$ when
$e$\ has type $\tau$ in $\E$ under the constraints $C$ and unifier $\unif$
with a usage environment $\Sv$. $\bm\E$ and $\bm e$ are the input parameters of our
inference algorithm.
Unlike in the syntax-directed version, $\E$ only regular and type bindings.
Suspended and borrow bindings can only be present in $\Sv$.
We revisit some of the syntax-directed rules presented in \cref{sdtyping}
to highlight the novelties
of our inference algorithm and the differences with the syntax-directed
system.
As before, the complete type inference rules are shown in \cref{appendix:infer}.

@@

As per tradition in ML languages, generalization is performed
on let-bindings.
Following HM(X), we note $(C_\schm, \schm) = \generalize{C}{\E}{\tau}$
the pair of a constraint and a scheme resulting from
generalization. The definition is provided in \cref{rule:infer:let}.
The type scheme $\schm$ is created by quantifying over all the appropriate
free variables and the current constraints.
The generated constraint $C_\schm$ uses a new projection operator,
noted $\Cproj{\tvar}{D}$, which
allow the creation of local variables inside the constraints.
This allows us to encapsulate all the quantified variables in the global constraints.
It also reflects the fact that there
should exist at least one solution for $C$ for the scheme to be valid.
\citet{DBLP:journals/tapos/OderskySW99} give a detailed account
on the projection operators in HM inference.

@@

Let $\CL$ be the set of constraints in such a lattice $\mathcal L$.
The full grammar of constraints is shown in \cref{grammar:constraint}.
Constraints are made of kind inequalities, conjunctions and
projections, as shown previously, along with type unification
constraints. Since types might contain kinds (for instance, on the arrows),
type unification is oriented and noted $\leq$.
For simplicity, we consider all type constructors
invariant in their parameters
and define $\Ceq{\tau}{\tau'}$ as $\Cleq{\tau}{\tau'} \wedge \Cleq{\tau'}{\tau}$.

As before, entailment is noted $\entail{C}{D}$,
where $D$ is a consequence of the constraints $C$.
We say that $C$ and $D$ are equivalent, noted $C \equivC D$,
when $\entail{C}{D}$ and $\entail{D}{C}$.
The base rules for entailment are given in \cref{rules:entail}.
The rules for classic Herbrand unification and
projection operators are omitted for brevity (see
\citep{DBLP:journals/tapos/OderskySW99} for
details).

@@

We define the set of solved formed
$\mathcal S$ as the set of constraints $\CL$ quantified by $\equivC$.
We will show later that such constraints are in fact only composed of
kind inequalities, and thus correspond to the syntactic constraints
used in type and kind schemes.
%
We now define our normalization procedure, noted $\normalize{C_0}{\unif_0}$ where
$C_0\in \CL$ is a set of constraints and $\unif_0$ is a substitution.
It returns a constraint $C \in \mathcal S$ in
solved form and a unifier $\unif$.
The main idea of the algorithm is to first remove all the type equalities
by using regular Herbrand unification. After that, we only have
a set of inequalities among kinds, which we can consider as a relation.
We can then saturate the relation,
unify all kinds that are in the same equivalence classes to obtain
a most general unifier on kind variables,
remove all existentially quantified variables and
then minimize back the relation and apply various
simplification rules to make the resulting type easier to understand to users.

@@

  We note $\mathcal R$ the relation $\Cleq{k_j}{k'_j}_j$,
  $\mathcal G$ the underlying directed graph and $V$ its vertices.

@@

\item For each strongly connected component in $\mathcal G$, unify all its vertices and replace it by a representative.
  We note $\unif_k$ the substitution that replaces a kind variable by
  its representative.
  The representative of a strongly connected component $g$ can be determined as follows:
  \begin{itemize}
  \item If $g$ does not contain any constant, then the representative
    is a fresh kind variable.
  \item If $g$ contains exactly one constant, it is the representative.
  \item Otherwise, the initial constraint $C_0$ is not satisfiable.
  \end{itemize}
  Note that this step will also detect all unsatisfiable constraints.
\item Take the transitive closure of $\mathcal R$.
\item Remove all the vertices corresponding to the kind variables $\kvar_i$
  that are existentially quantified in $C^k$.
\item Take the transitive reduction of $\mathcal R$.
\item Remove the extremums of $\mathcal L$ and the edges of $(\leq)$
  from $\mathcal R$.
\item Return $C = \left\{ k \leq k' \mid k \operatorname{\mathcal R}k' \right\}$
  and $\unif =  \unif_\tau \meet \unif_k$.
\end{enumerate}

@@

\begin{definition}[Flattened Environment]
A flattened environment,
noted $\Eflat\E$, is the environment
where all the binders are replace by normal ones. More formally:
\begin{align*}
  \Eflat\E
  &= \left\{ \bvar{x}{\tau} \mid
    \bvar{x}{\tau}\in\E
    \vee \bvar{\borrow{x}}{\borrowty{k}{\tau}}\in\E
    \vee \svar{x}{\tau}^n\in\E
    \right\}\\
  &\cup \left\{ \bvar{\tvar}{k} \mid \bvar{\tvar}{k}\in\E \right\}
\end{align*}
\end{definition}

>>>>

Unlike the syntax-directed version of our type system, knowing which elements
are input and output or our inference judgement is critical. In the rest
of this section, when presenting a new judgement,
we write input parameters in \textbf{bold}. The rest will then be
output parameters.

@@

% One novelty of our inference  judgement now
% returns a type environment, which we call ``usage environment'' and commonly
% write $\Sv$, to summarize how variables and borrows are used inside
% the expression.

In order to determine if a variable is used in an affine manner, we must track
its uses and the associated kinds. For instance, in the expression
$(x,x)$, $x$ is used twice. If $x$ is of type $\tau$, which is of kind $k$,
we must add the constraint $\Cleq{k}{\kun}$.
%
In order to infer such constraints, our inference judgement will not only
take an environment as parameter but also return an environment which
we call ``usage environment'' and which summarizes how variables and borrows
are used. Usage environments follow the exact same grammar
as normal environments. In order to distinguish them more easily,
we denote them by $\Sv$.

@@

We write $\inferW{\Sv}{(C,\unif)}{\bm{\E}}{\bm{e}}{\tau}$ when
$e$\ has type $\tau$ in $\E$ under the constraints $C$ and unifier $\unif$
with a usage environment $\Sv$. $\bm\E$ and $\bm e$ are the input parameters of our
inference algorithm.
Unlike in the syntax-directed version, $\E$ only regular and type bindings.
Suspended and borrow bindings can only be present in $\Sv$.
We revisit some of the syntax-directed rules presented in \cref{sdtyping}
to highlight the novelties
of our inference algorithm and the differences with the syntax-directed
system.
As before, the complete type inference rules are shown in \cref{appendix:infer}.

@@

As per tradition in ML languages, generalization is performed
on let-bindings.
Following HM(X), we write $(C_\schm, \schm) = \generalize{C}{\E}{\tau}$
for the pair of a constraint and a scheme resulting from
generalization. The definition is provided in \cref{rule:infer:let}.
The type scheme $\schm$ is created by quantifying over all the appropriate
free variables and the current constraints.
The generated constraint $C_\schm$ uses a new projection operator,
$\Cproj{\tvar}{D}$, which
allow the creation of local variables inside the constraints.
This allows us to encapsulate all the quantified variables in the global constraints.
It also reflects the fact that there
should exist at least one solution for $C$ for the scheme to be valid.
\citet{DBLP:journals/tapos/OderskySW99} give a detailed account
on the projection operators in HM inference.

@@

Let $\CL$ be the set of constraints in such a lattice $\mathcal L$.
The full grammar of constraints is shown in \cref{grammar:constraint}.
Constraints are made of kind inequalities, conjunctions and
projections, as shown previously, along with type unification
constraints. Since types might contain kinds (for instance, on the arrows),
type unification is oriented and written as $\leq$.
For simplicity, we consider all type constructors
invariant in their parameters
and define $\Ceq{\tau}{\tau'}$ as $\Cleq{\tau}{\tau'} \wedge \Cleq{\tau'}{\tau}$.

As before, entailment is denoted by $\entail{C}{D}$,
where $D$ is a consequence of the constraints $C$.
We say that $C$ and $D$ are equivalent, $C \equivC D$,
when $\entail{C}{D}$ and $\entail{D}{C}$.
The base rules for entailment are given in \cref{rules:entail}.
The rules for classic Herbrand unification and
projection operators are omitted for brevity (see
\citep{DBLP:journals/tapos/OderskySW99} for
details).

@@

We define the set of solved formed
$\mathcal S$ as the set of constraints $\CL$ quantified by $\equivC$.
We will show later that such constraints are in fact only composed of
kind inequalities, and thus correspond to the syntactic constraints
used in type and kind schemes.
%
We now define our normalization procedure $\normalize{C_0}{\unif_0}$, where
$C_0\in \CL$ is a set of constraints and $\unif_0$ is a substitution.
It returns a constraint $C \in \mathcal S$ in
solved form and a unifier $\unif$.
The main idea of the algorithm is to first remove all the type equalities
by using regular Herbrand unification. After that, we only have
a set of inequalities among kinds, which we can consider as a relation.
We can then saturate the relation,
unify all kinds that are in the same equivalence classes to obtain
a most general unifier on kind variables,
remove all existentially quantified variables and
then minimize back the relation and apply various
simplification rules to make the resulting type easier to understand to users.

@@

  We write $\mathcal R$ for the relation $\Cleq{k_j}{k'_j}_j$,
  $\mathcal G$ the underlying directed graph and $V$ its vertices.

@@

\item For each strongly connected component in $\mathcal G$, unify all its vertices and replace it by a representative.
  We write $\unif_k$ for the substitution that replaces a kind variable by
  its representative.
  The representative of a strongly connected component $g$ can be determined as follows:
  \begin{itemize}
  \item If $g$ does not contain any constant, then the representative
    is a fresh kind variable.
  \item If $g$ contains exactly one constant, it is the representative.
  \item Otherwise, the initial constraint $C_0$ is not satisfiable.
  \end{itemize}
  Note that this step will also detect all unsatisfiable constraints.
\item Take the transitive closure of $\mathcal R$.
\item Remove all the vertices corresponding to the kind variables $\kvar_i$
  that are existentially quantified in $C^k$.
\item Take the transitive reduction of $\mathcal R$.
\item Remove the extremums of $\mathcal L$ and the edges of $(\leq)$
  from $\mathcal R$.
\item Return $C = \left\{ k \leq k' \mid k \operatorname{\mathcal R}k' \right\}$
  and $\unif =  \unif_\tau \meet \unif_k$.
\end{enumerate}

@@

\begin{definition}[Flattened Environment]
A flattened environment,
written as $\Eflat\E$, is the environment
where all the binders are replace by normal ones. More formally:
\begin{align*}
  \Eflat\E
  &= \left\{ \bvar{x}{\tau} \mid
    \bvar{x}{\tau}\in\E
    \vee \bvar{\borrow{x}}{\borrowty{k}{\tau}}\in\E
    \vee \svar{x}{\tau}^n\in\E
    \right\}\\
  &\cup \left\{ \bvar{\tvar}{k} \mid \bvar{\tvar}{k}\in\E \right\}
\end{align*}
\end{definition}

### 2019-11-11 kindly/sdtyping/extra.tex
Reformat the annex a bit.
Start improving the inference section.
Fixed uses of 'note' instead of 'denote' and various typos.

So far, all \lang programs have been fully annotated with regions information.
We now show how to infer these regions annotations based on
optionally-annotated programs.
First, we extend the region annotation to $\region{S}{E}$ where $S$ is
a map from variables to borrow indicator $b$. This annotation, defined below, is equivalent to nested
region annotations for each individual variable.
%
\begin{align*}
  \region{\Sone x\BORROW;S}{e} &= \region{\Sone x\BORROW}{\region{S}{e}}& \region{\emptyset}{e} &= e\\
\end{align*}
%
\begin{figure*}[b]
  \centering
  \input{region-annot}
  \caption{Automatic region annotation --- $\RannotT{e}{e'}$}
  \label{fig:region-annotation}
\end{figure*}
%
\Cref{fig:region-annotation} define a rewriting relation $\RannotT{e}{e'}$
which indicates that an optionally annotated term $e$ can be rewritten
as a fully annotated term $e'$.
Through the rule \textsc{Rewrite-Top}, this is defined
in term of an inductively defined relation
$\Rannot{e}{e'}{S}$ where $n$ is the current nesting and $S$ is a set of
variable that are not yet enclosed in a region.
The base cases are constants, variables and borrows.
The general idea is to start from the leaves of the syntax tree, create a
region for each borrow, and enlarge the region as much as possible.
This is implemented by a depth-first walk of the syntax
tree which collects each variable that has a corresponding borrow.
At each step, it rewrites the inner subterms,
consider which borrow must be enclosed by a region now, and
return the others for later enclosing. Binders force immediate
enclosing of the bound variables, as demonstrated in rule \textsc{Rewrite-Lam}.
For nodes with multiple children, we
use a scope merge operator to decide if regions should be placed and where.
This is shown in rule \textsc{Rewrite-Pair}.
The merge operator, written $\getBorrows{B_l}{B_r}{(S_l,S,S_r)}$, takes
the sets $B_l$ and $B_r$ returned by rewriting the subterms
and returns three sets: $S_l$ and $S_r$ indicates the variables
that should be immediately enclosed by a region on the left and right
subterms and $S$ indicates the set of the yet-to-be-enclosed variables.
As an example, the rule \textsc{AnnotRegion-MutLeft} is applied
when there is an immutable borrow and a mutable borrow. In that case, a
region is created to enclose the immutable borrow, while the mutable
borrow is left to be closed later. This is coherent with the rules
for environment splitting and suspended bindings from \cref{sdtyping}.
%
Explicitly annotated regions are handled specially through
rule \textsc{Rewrite-Region}. In that case, we assume that all inner
borrows should be enclosed immediately.

@@

Our system is defined in \cref{infer:solving}.
As before, we note $\CL$ our constraint system for an arbitrary
total bounded lattice $(\mathcal L, \lk_\Lat)$.
Entailment is noted $\entail{C}{D}$, where $D$ is a consequence of the
constraints $C$.
We say that $C$ and $D$ are equivalent, noted $C \equivC D$,
when $\entail{C}{D}$ and $\entail{D}{C}$.
Following HM(X), we define a cylindric constraint system as
a system which respects the following properties for any variable $x$:
\begin{align*}
  \entail{C&}{\Cproj{x}{C}}
  &\entail{C}{D} &\implies \entail{\Cproj{x}C}{\Cproj{x}D}\\
  \Cproj{x}{(C\wedge \Cproj{x}D)}&\equivC \Cproj{x}{C} \wedge \Cproj{x}D
  & \Cproj{x}{\Cproj{y}D}&\equivC \Cproj{y}{\Cproj{x}D}
\end{align*}

Furthermore, we define a term rewriting system as a system where, for every
types $\tau$,$\tau'$, there exists an equality predicates $\Ceq{\tau}{\tau'}$
which is symmetric, reflexive, transitive, stable under substitution and such that,
for any predicate $P$:
\begin{align*}
  \entail{\Ceq{x}{y}\wedge \Cproj{x}C \wedge \Ceq{x}{y}&}{C}
  &\subst{x}{\tau}{P} &\equivC \Cproj{x}P \wedge \Ceq{x}{\tau}
                        \text{ where } x\notin\fv{\tau}
\end{align*}
$\CL$ is defined as the smallest cylindric term constraint system that
satisfies the axiom shown in \cref{rules:entail}.
We define  $\Ceq{\tau}{\tau'}$ as $\Cleq{\tau}{\tau'} \wedge \Cleq{\tau'}{\tau}$.
%
We note $\SC$ the set of solved forms
which can be used inside type and kind schemes.
We define $\SC$ as $\CL$ quotiented by the relation $\equivC$.
%
The normalization function is described in \cref{infer:solving}.

@@

  We note $(C^k,\unif_\tau)$ the result of
  the modified Herbrand unification on $(D,\phi)$ in step (1).
  Herbrand unification computes the most general
  unifier. Our modified Herbrand unification only output additional
  kind constraints for kind on the arrows and does not change
  the result of the unification. Thus, we have
  $\phi_\tau\leq\unif_\tau$,
  $C^k \equivC \unif_\tau D$ and
  $\unif_\tau C^k = C^k$.

@@

  We note $P(C_a)$ if for all $C = (k, k)'$
  such that $\entail{C_a}{C}$ and $\nvdash_eC$,
  we have $C \in {\mathcal R}_a$.

  Let us note $C_1'$ and $C_2'$ the constraints after step (4). $P(C_1')$ and
  $P(C_2')$ hold. Indeed, since $C_1'$ and $C_2'$ are only composed
  of existential quantifications and kind inequalities, the only rules
  that applies are transitivity and lattice inequalities.
  After step (2) and (5), the associated relations are fully saturated for these
  two rules, hence all inequalities that can be deduced from $C_a'$ are already
  present in the relation.

@@

  Let us note $C''_a$ for $a\in\{1,2\}$ the constraints after step (5).
  Since there are no more existential variables,
  we have $C''_a=(k_i,k'_i)_i=\mathcal R''_a$.
  For any $C=(k,k')$ such that $\entail{}{C}$ and $\entail{C''_a}{C}$,
  then $C \in (\leq) \subset {\mathcal R}''_a$.
  Indeed, the only trivial inequalities in our system are equalities of the form
  $(\kvar, \kvar)$, which were removed in step (4) and the lattice inequalities.

@@

\subsection{Kinding}
%
\begin{figure}[tb]
  \centering
  \input{sdtyping/sd-kinds}
  \caption{Syntax-directed kinding rule --
    $\inferSK{C}{\E}{\tau}{k}$}
  \label{rules:sd-kinding}
\end{figure}

We note $\inferSK{C}{\E}{\tau}{k}$
when $\tau$ has kind $k$ in environment $\E$ under constraints $C$.
The rules are shown in \cref{rules:sd-kinding}.
Kinds and types follow a small calculus with variables ($\tvar$,\dots),
functions (type constructors $\T{t}$), application ($\tapp{t}{\Multi{\tau}}$)
and primitives such as types for arrows ($\tau\tarr{k}\tau'$) and
borrows ($\borrowty{k}{\tau}$).
Kind checking can thus be done in a fairly straightforward, syntax-directed
fashion by simply following
the syntax of the types. Kind arrows can only appear when looking
up the kind scheme of a type constructor $\T t$. Kind arrows are forbidden
in any other contexts.

@@

\cref{fig:contraints-environments-types} defines
the rewriting relation on environment constraints,
noted $\Cleq\E k \Crewrite  C$ which rewrites a constraint of the form
$\Cleq\E k$ into $C$. It proceeds by iterating over the environment
and expending the constraints for each binding.
Suspended bindings are rejected ({\sc ConstrSusp}).
Borrow bindings directly use the annotated kind ({\sc ConstrBorrow}).
Other bindings uses the underlying type scheme ({\sc ConstrBinding}).
Type schemes are constrained by first inferring the kind, and then
emitting the constraint ({\sc ConstrSD} and {\sc ConstrI}).

\begin{figure}[!tb]
  \input{sdtyping/sd-splitting}
  \caption{Splitting --- environments $\lsplit
    C\E\E\E$; binders $\bsplit Cbbb$}
  \label{fig:sd-splitting}
  \input{sdtyping/sd-borrowing}
  \caption{Borrowing --- environments $\lregion
    C{x}\E\E$; binders $\bregion C{x}bb$}
  \label{fig:sd-borrowing}
  \input{sdtyping/sd-environments}
  \caption{Rewriting constraints on environments --- $\Cleq\E k \Crewrite  C$}
  \label{fig:contraints-environments-types}
\end{figure}

>>>>

So far, all \lang programs have been fully annotated with regions information.
We now show how to infer these regions annotations based on
optionally-annotated programs.
First, we extend the region annotation to $\region{S}{E}$ where $S$ is
a map from variables to borrow indicator $b$. This annotation, defined below, is equivalent to nested
region annotations for each individual variable.
%
\begin{align*}
  \region{\Sone x\BORROW;S}{e} &= \region{\Sone x\BORROW}{\region{S}{e}}& \region{\emptyset}{e} &= e\\
\end{align*}
%
\begin{figure*}[!tb]
  \centering
  \input{region-annot}
  \caption{Automatic region annotation --- $\RannotT{e}{e'}$}
  \label{fig:region-annotation}
\end{figure*}
%
\Cref{fig:region-annotation} define a rewriting relation $\RannotT{e}{e'}$
which indicates that an optionally annotated term $e$ can be rewritten
as a fully annotated term $e'$.
Through the rule \textsc{Rewrite-Top}, this is defined
in term of an inductively defined relation
$\Rannot{e}{e'}{S}$ where $n$ is the current nesting and $S$ is a set of
variable that are not yet enclosed in a region.
The base cases are constants, variables and borrows.
The general idea is to start from the leaves of the syntax tree, create a
region for each borrow, and enlarge the region as much as possible.
This is implemented by a depth-first walk of the syntax
tree which collects each variable that has a corresponding borrow.
At each step, it rewrites the inner subterms,
consider which borrow must be enclosed by a region now, and
return the others for later enclosing. Binders force immediate
enclosing of the bound variables, as demonstrated in rule \textsc{Rewrite-Lam}.
For nodes with multiple children, we
use a scope merge operator to decide if regions should be placed and where.
This is shown in rule \textsc{Rewrite-Pair}.
The merge operator, written $\getBorrows{B_l}{B_r}{(S_l,S,S_r)}$, takes
the sets $B_l$ and $B_r$ returned by rewriting the subterms
and returns three sets: $S_l$ and $S_r$ indicates the variables
that should be immediately enclosed by a region on the left and right
subterms and $S$ indicates the set of the yet-to-be-enclosed variables.
As an example, the rule \textsc{AnnotRegion-MutLeft} is applied
when there is an immutable borrow and a mutable borrow. In that case, a
region is created to enclose the immutable borrow, while the mutable
borrow is left to be closed later. This is coherent with the rules
for environment splitting and suspended bindings from \cref{sdtyping}.
%
Explicitly annotated regions are handled specially through
rule \textsc{Rewrite-Region}. In that case, we assume that all inner
borrows should be enclosed immediately.

@@

Our system is defined in \cref{infer:solving}.
As before, we write $\CL$ for our constraint system for an arbitrary
total bounded lattice $(\mathcal L, \lk_\Lat)$.
Entailment is written $\entail{C}{D}$, where $D$ is a consequence of the
constraints $C$.
We say that $C$ and $D$ are equivalent, written as $C \equivC D$,
when $\entail{C}{D}$ and $\entail{D}{C}$.
Following HM(X), we define a cylindric constraint system as
a system which respects the following properties for any variable $x$:
\begin{align*}
  \entail{C&}{\Cproj{x}{C}}
  &\entail{C}{D} &\implies \entail{\Cproj{x}C}{\Cproj{x}D}\\
  \Cproj{x}{(C\wedge \Cproj{x}D)}&\equivC \Cproj{x}{C} \wedge \Cproj{x}D
  & \Cproj{x}{\Cproj{y}D}&\equivC \Cproj{y}{\Cproj{x}D}
\end{align*}

Furthermore, we define a term rewriting system as a system where, for every
types $\tau$,$\tau'$, there exists an equality predicates $\Ceq{\tau}{\tau'}$
which is symmetric, reflexive, transitive, stable under substitution and such that,
for any predicate $P$:
\begin{align*}
  \entail{\Ceq{x}{y}\wedge \Cproj{x}C \wedge \Ceq{x}{y}&}{C}
  &\subst{x}{\tau}{P} &\equivC \Cproj{x}P \wedge \Ceq{x}{\tau}
                        \text{ where } x\notin\fv{\tau}
\end{align*}
$\CL$ is defined as the smallest cylindric term constraint system that
satisfies the axiom shown in \cref{rules:entail}.
We define  $\Ceq{\tau}{\tau'}$ as $\Cleq{\tau}{\tau'} \wedge \Cleq{\tau'}{\tau}$.
%
We write $\SC$ for the set of solved forms,
which can be used inside type and kind schemes.
We define $\SC$ as $\CL$ quotiented by the relation $\equivC$.
%
The normalization function is described in \cref{infer:solving}.

@@

  We write $(C^k,\unif_\tau)$ for the result of
  the modified Herbrand unification on $(D,\phi)$ in step (1).
  Herbrand unification computes the most general
  unifier. Our modified Herbrand unification only output additional
  kind constraints for kind on the arrows and does not change
  the result of the unification. Thus, we have
  $\phi_\tau\leq\unif_\tau$,
  $C^k \equivC \unif_\tau D$ and
  $\unif_\tau C^k = C^k$.

@@

  We write $P(C_a)$ if for all $C = (k, k)'$
  such that $\entail{C_a}{C}$ and $\nvdash_eC$,
  we have $C \in {\mathcal R}_a$.

  Let us write $C_1'$ and $C_2'$ for the constraints after step (4). $P(C_1')$ and
  $P(C_2')$ hold. Indeed, since $C_1'$ and $C_2'$ are only composed
  of existential quantifications and kind inequalities, the only rules
  that applies are transitivity and lattice inequalities.
  After step (2) and (5), the associated relations are fully saturated for these
  two rules, hence all inequalities that can be deduced from $C_a'$ are already
  present in the relation.

@@

  Let us write $C''_a$ for $a\in\{1,2\}$ the constraints after step (5).
  Since there are no more existential variables,
  we have $C''_a=(k_i,k'_i)_i=\mathcal R''_a$.
  For any $C=(k,k')$ such that $\entail{}{C}$ and $\entail{C''_a}{C}$,
  then $C \in (\leq) \subset {\mathcal R}''_a$.
  Indeed, the only trivial inequalities in our system are equalities of the form
  $(\kvar, \kvar)$, which were removed in step (4) and the lattice inequalities.

@@

\subsection{Kinding}
%
\begin{figure}[bt]
  \centering
  \input{sdtyping/sd-kinds}
  \caption{Syntax-directed kinding rule --
    $\inferSK{C}{\E}{\tau}{k}$}
  \label{rules:sd-kinding}
\end{figure}

We write $\inferSK{C}{\E}{\tau}{k}$
if $\tau$ has kind $k$ in environment $\E$ under constraints $C$.
The rules are shown in \cref{rules:sd-kinding}.
Kinds and types follow a small calculus with variables ($\tvar$,\dots),
functions (type constructors $\T{t}$), application ($\tapp{t}{\Multi{\tau}}$)
and primitives such as types for arrows ($\tau\tarr{k}\tau'$) and
borrows ($\borrowty{k}{\tau}$).
Kind checking can thus be done in a fairly straightforward, syntax-directed
fashion by simply following
the syntax of the types. Kind arrows can only appear when looking
up the kind scheme of a type constructor $\T t$. Kind arrows are forbidden
in any other contexts.

@@

\cref{fig:contraints-environments-types} defines
the rewriting relation on environment constraints,
$\Cleq\E k \Crewrite C$, which rewrites a constraint of the form
$\Cleq\E k$ into $C$. It proceeds by iterating over the environment
and expending the constraints for each binding.
Suspended bindings are rejected ({\sc ConstrSusp}).
Borrow bindings directly use the annotated kind ({\sc ConstrBorrow}).
Other bindings uses the underlying type scheme ({\sc ConstrBinding}).
Type schemes are constrained by first inferring the kind, and then
emitting the constraint ({\sc ConstrSD} and {\sc ConstrI}).

\begin{figure*}[!tb]
  \input{sdtyping/sd-splitting}
  \caption{Splitting ---
    environments $\lsplit C\E{\E_l}{\E_r}$;
    inference $\bsplit C\E{\bm\E_l}{\bm\E_r}$;
    binders $\bsplit Cb{\bm b_r}{\bm b_l}$}
  \label{fig:sd-splitting}
  \input{sdtyping/sd-borrowing}
  \caption{Borrowing ---
    environments $\lregion[n]{C}{x}{\E}{\E'}$;
    inference $\bregion[\bm n]{C}{\bm x}{\E}{\bm\E'}$;
    binders $\bregion[\bm n]{C}{\bm x}{b}{\bm b'}$}
  \label{fig:sd-borrowing}
  \input{sdtyping/sd-environments}
  \caption{Rewriting constraints on environments --- $\bm{\Cleq{\E}{k}}\Crewrite  C$}
  \label{fig:contraints-environments-types}
\end{figure*}
